Covariates analysis of MSBB RNASeq data

### 2016-01-26 code/Rmd/MSSM_RNASeq.Rmd
Added differential expression analysis to MSSM RNASeq markdown

tree = hclust(as.dist(t(RESIDUAL.GENE_EXPRESSION)))
cols = WGCNA::labels2colors(COVARIATES.tmp);
tmp = gsub('\\..*','',cols)
WGCNA::plotDendroAndColors(tree, 
                           colors = tmp, 
                           dendroLabels = FALSE, 
                           abHeight = 0.80, 
                           main = "Sample dendrogram",
                           groupLabels = colnames(COVARIATES.tmp))
```

```{r summary, results='asis', echo=FALSE}
writeLines('Number of samples in each category are')
tmp = as.data.frame(summary(COVARIATES$TISSUE.Dx))
colnames(tmp) = 'No. of samples'
kable(tmp)
```

@@

# Covariates
covar = rownameToFirstColumn(COVARIATES, 'SampleId')
write.table(covar, file='covariates.tsv', sep = '\t', row.names=F, quote=F)
obj = File('covariates.tsv', 
           name = 'Covariates', 
           parentId = CODE$properties$id)
covar_obj = synStore(obj,
               used = ALL_USED_IDs, 
               activityName = activityName, 
               executed = thisFile, 
               activityDescription = activityDescription)
```
|  *Results*            |  *SynapseID*                    |
|  -------              |   ---------                     |
|  Covariates           |  `r covar_obj$properties$id`    |
|  Residual Expression  |  `r expr_obj$properties$id`     |

>>>>

tree = hclust(as.dist(t(RESIDUAL.GENE_EXPRESSION)))
cols = WGCNA::labels2colors(COVARIATES.tmp);
tmp = gsub('\\..*','',cols)
WGCNA::plotDendroAndColors(tree, 
                           colors = tmp, 
                           dendroLabels = FALSE, 
                           abHeight = 0.80, 
                           main = "Sample dendrogram",
                           groupLabels = colnames(COVARIATES.tmp))
```
```{r get.gene.symbols, cache=TRUE}
# Define biomart object
mart <- useMart(biomart = "ensembl", dataset = "hsapiens_gene_ensembl")

# Query biomart
Ensemble2HGNC <- getBM(attributes = c("ensembl_gene_id", "hgnc_symbol"),
                       filters = "ensembl_gene_id", values = rownames(RESIDUAL.GENE_EXPRESSION),
                       mart = mart)
RESIDUAL.GENE_EXPRESSION <- left_join(rownameToFirstColumn(RESIDUAL.GENE_EXPRESSION, "ensembl_gene_id"),
                                      Ensemble2HGNC) %>%
  dplyr::select(ensembl_gene_id, hgnc_symbol, one_of(rownames(COVARIATES)))
```
```{r summary, results='asis', echo=FALSE}
writeLines('Number of samples in each category are')
tmp = as.data.frame(summary(COVARIATES$TISSUE.Dx))
colnames(tmp) = 'No. of samples'
kable(tmp)
```
### Differential expression analysis
Genes that are differentially expressed at an FDR <= 0.05 are
```{r diffExp}
# Fit contrast
contrast = makeContrasts(contrasts=c("TISSUE.DxBM_10.Control-TISSUE.DxBM_10.AD",
                                     "TISSUE.DxBM_22.Control-TISSUE.DxBM_22.AD",
                                     "TISSUE.DxBM_36.Control-TISSUE.DxBM_36.AD"), 
                         levels = colnames(VOOM.ADJUSTED.FIT$coefficients))
FIT = contrasts.fit(VOOM.ADJUSTED.FIT, contrasts=contrast)
FIT = eBayes(FIT)

# Get differnetial expression
DE = list()
p = list()
for (i in 1:3){
  DE[[i]] = topTable(FIT, coef=i, number = dim(VOOM.ADJUSTED.GENE_EXPRESSION$E)[1])
  DE[[i]] = rownameToFirstColumn(DE[[i]], "ensembl_gene_id") %>%
    left_join(Ensemble2HGNC)
  p[[i]] <- ggplot(DE[[i]], aes(x = logFC, y = -log10(adj.P.Val))) + geom_point()
  p[[i]] <- p[[i]] + ggtitle(gsub('TISSUE.Dx','',colnames(contrast)[i])) + xlim(c(-1.5,1.5)) + ylim(c(0,6))
  p[[i]] <- p[[i]] + geom_hline(yintercept = -log10(0.05), color = 'red')
}
multiplot(plotlist=p, cols = 3)
names(DE) = gsub('TISSUE.Dx','',colnames(contrast))

DE = DE %>% ldply(.id = 'Comparison')

tmp = DE %>%
  dplyr::filter(adj.P.Val <= 0.05) %>%
  dplyr::select(Comparison, hgnc_symbol) %>%
  group_by(Comparison) %>%
  summarise(nDiffExp = length(unique(hgnc_symbol)), 
            DiffExpGenes = paste(unique(hgnc_symbol), collapse = ','))

kable(tmp)
```

@@

# Covariates
covar = rownameToFirstColumn(COVARIATES, 'SampleId')
write.table(covar, file='covariates.tsv', sep = '\t', row.names=F, quote=F)
obj = File('covariates.tsv', 
           name = 'Covariates', 
           parentId = CODE$properties$id)
covar_obj = synStore(obj,
               used = ALL_USED_IDs, 
               activityName = activityName, 
               executed = thisFile, 
               activityDescription = activityDescription)

# Store differential expression results
write.table(DE, file = 'diffExpression.tsv', sep = '\t', row.names=F, quote=F)
DEXP_OBJ = File('diffExpression.tsv', name = 'Differential Expression Analysis', parentId = CODE$properties$id)
DEXP_OBJ = synStore(DEXP_OBJ, used = ALL_USED_IDs, activityName = activityName, 
                   executed = thisFile, activityDescription = activityDescription)
```
|  *Results*            |  *SynapseID*                    |
|  -------              |   ---------                     |
|  Covariates           |  `r covar_obj$properties$id`    |
|  Residual Expression  |  `r expr_obj$properties$id`     |
|  Differential Expression  |  `r DEXP_OBJ$properties$id`      |
